---
title: "Computation time and features"
author: "Benjamin Christoffersen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Computation time and features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
options(width = 80)
knitr::opts_chunk$set(collapse = TRUE, comment = "#R>", error = FALSE)
```

This vignette shows comparisons in terms of computation with other packages and 
alternative `base` R solutions. Specifially, we will make comparissons with the 
`roll` package and `zoo` package. It should be stressed though that the other 
solutions do additional things than this package does. E.g., there is not 
performed any rank test in the function in this package. We start by showing 
the comparisons of computation time and then we show different options. R 
versions of the C++ function are shown to make it clear what the output is.

# Comparisons

```{r def_roll_funcs, echo = FALSE, message=FALSE}
#####
# Pure R version of package function
library(SamplerCompare) # for LINPACK `chdd` and `chud`
roll_regress_R <- function(X, y, width){
  n <- nrow(X)
  p <- ncol(X)
  out <- matrix(NA_real_, p, n)

  is_first <- TRUE
  for(i in width:n){
    if(is_first){
      is_first <- FALSE
      qr. <- qr(X[1:width, ])
      R <- qr.R(qr.)

      # Use X^T for the rest
      X <- t(X)

      XtY <- drop(tcrossprod(y[1:width], X[, 1:width]))
    } else {
      x_new <- X[, i]
      x_old <- X[, i - width]

      # update R 
      R <- .Fortran(
        "dchud", R, p, p, x_new, 0., 0L, 0L, 
        0., 0., numeric(p), numeric(p), 
        PACKAGE = "SamplerCompare")[[1]]

      # downdate R
      R <- .Fortran(
        "dchdd", R, p, p, x_old, 0., 0L, 0L, 
        0., 0., numeric(p), numeric(p), integer(1),
        PACKAGE = "SamplerCompare")[[1]]

      # update XtY
      XtY <- XtY + y[i] * x_new - y[i - width] * x_old
    }

    coef. <- .Internal(backsolve(R, XtY, p, TRUE, TRUE))
    coef. <- .Internal(backsolve(R, coef., p, TRUE, FALSE))

    out[, i] <- coef.
  }

  t(out)
}

#####
# simple R version
roll_regress_R_for_loop <- function(X, y, width){
  n <- nrow(X)
  p <- ncol(X)
  out <- matrix(NA_real_, n, p)

  for(i in width:n){
    idx <- (i - width + 1L):i
    out[i, ] <- lm.fit(X[idx, , drop = FALSE], y[idx])$coefficients
  }

  out
}

#####
# zoo version
.zoo_inner <- function(Z) {
  fit <- lm.fit(Z[, -1, drop = FALSE], Z[, 1])
  fit$coefficients
}
library(zoo)
roll_regress_zoo <- function(x, y, width){
  Z <- cbind(y, X)
  rollapply(Z, width, FUN = .zoo_inner,  
            by.column = FALSE,  align = "right", fill = NA_real_)
}

#####
# roll lm
library(roll)
roll_lm_func <- function(x, y ,width)
  roll_lm(X, matrix(y, ncol = 1), wdth, intercept = FALSE)$coefficients[, -1L]

# use one thread as other methods are easily to compute in parallel too
RcppParallel::setThreadOptions(numThreads = 1)

# compile functions
library(compiler)
roll_regress_R          <- cmpfun(roll_regress_R)
roll_regress_R_for_loop <- cmpfun(roll_regress_R_for_loop)
.zoo_inner              <- cmpfun(.zoo_inner)
roll_regress_zoo        <- cmpfun(roll_regress_zoo)
roll_lm_func            <- cmpfun(roll_lm_func)
```

We start by Simulating the data

```{r sim_data}
set.seed(101)
n <- 10000
p <- 6
wdth = 50
X <- matrix(rnorm(p * n), n, p)
y <- drop(X %*% runif(p)) + rnorm(n)
df <- data.frame(y, X)
frm <- eval(parse(text = paste0(
  "formula(y ~ -1 + ", paste0("X", 1:p, collapse = " + "), ")")))
```

Then we check that the functions give the same (the function definitions are at
the end of this document)

```{r roll_gives_same}
library(rollRegres)
all.equal(roll_regress_R(X, y, wdth), roll_regres.fit(X, y, wdth), 
          check.attributes = FALSE)
all.equal(roll_regress_R(X, y, wdth), roll_regres(frm, df, wdth), 
          check.attributes = FALSE)
all.equal(roll_regress_R(X, y, wdth), roll_regress_zoo(X, y, wdth), 
          check.attributes = FALSE)
all.equal(roll_regress_R(X, y, wdth), roll_regress_R_for_loop(X, y, wdth), 
          check.attributes = FALSE)
all.equal(roll_regress_R(X, y, wdth), roll_lm_func(X, y, wdth), 
          check.attributes = FALSE)
```

and here then we compare the computation time

```{r benchmark_roll}
microbenchmark::microbenchmark(
  roll_regress            = roll_regres.fit(X, y, wdth),
  # this will be slower due to call to `model.matrix` and `model.frame`
  roll_regress_df         = roll_regres(frm, df, wdth),
  roll_regress_R          = roll_regress_R(X, y, wdth),
  roll_regress_zoo        = roll_regress_zoo(X, y, wdth),
  roll_regress_R_for_loop = roll_regress_R_for_loop(X, y, wdth),
  roll_lm = roll_lm_func(X, y, wdth),
  times = 5)

# here is the formula used above
frm
```

# Additional features

This section will cover some additional features.

## Expanding window

Here are expanding window regressions with additional output

```{r expanding_window}
#####
# simulate data

```

## Update in blocks
You can use the `grp` argument to make updates in block. E.g., here is an 
example with weekly data

```{r block_example}
#####
# simulate data
set.seed(68799379)
week <- as.integer(gl(25, 7))
n <- length(week)
p <- 2
X <- matrix(rnorm(p * n), n, p)
y <- drop(X %*% runif(p)) + rnorm(n)

#####
# use package function
pck_out <- roll_regres.fit(
  X, y, width = 10L, grp = week, 
  do_compute = c("sigmas", "r.squareds", "1_step_forecasts"))

#####
# make R-version
R_func <- function(X, y, width, grp){
  u_grp = unique(grp)
  n <- nrow(X)
  p <- ncol(X)
  out <- matrix(NA_real_, n, p)
  sigmas             <- rep(NA_real_, n)
  r.squared          <- rep(NA_real_, n)
  one_step_forecasts <- rep(NA_real_, n)

  start_val <- max(which(u_grp <= width))
  for(g in u_grp[start_val:length(u_grp)]){
    idx <- which(grp %in% (g - width + 1L):g)
    i <- which(grp == g)
    fit <- lm(y[idx] ~ -1 + X[idx, , drop = FALSE])
    out[i, ] <- sapply(fit$coefficients, rep, times = length(i))

    su <- summary(fit)
    sigmas[i] <- su$sigma

    ss1 <- sum((y[idx] - mean(y[idx]))^2)
    ss2 <- sum(fit$residuals^2)
    r.squared[i] <- 1 - ss2 / ss1

    if(g != max(grp)){
      next_g <- grp[min(which(grp > g))]
      next_g <- which(grp == next_g)
      one_step_forecasts[next_g] <- X[next_g, ] %*% fit$coefficients
    }
  }

  list(coef = out, sigmas = sigmas, r.squared = r.squared,
       one_step_forecasts = one_step_forecasts)
}

R_out <- R_func(X, y, 10L, grp = week)

#####
# gives the same
stopifnot(isTRUE(all.equal(R_out$coef              , pck_out$coefs)))
stopifnot(isTRUE(all.equal(R_out$sigmas            , pck_out$sigmas)))
stopifnot(isTRUE(all.equal(R_out$r.squared         , pck_out$r.squared)))
stopifnot(isTRUE(all.equal(R_out$one_step_forecasts, pck_out$one_step_forecasts)))
```

# Session info

```{r ses_info}
sessionInfo()
```

# Function definitions

Here are the function definitions

```{r def_roll_funcs, eval = FALSE}
```
