---
title: "Computation time and features"
author: "Benjamin Christoffersen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Computation time and features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
options(width = 80)
knitr::opts_chunk$set(collapse = TRUE, comment = "#R", error = FALSE)
```

This vignette shows comparisons in terms of computation time with other packages and 
alternative `base` R solutions. Specifically, we will make comparisons with the 
`roll` package and `zoo` package. It should be stressed though that the other 
solutions do additional things than this package does. E.g., there is not 
performed any rank test in the function in this package. We start by showing 
the comparisons of computation times and then we show different options. R 
versions of the functions are shown to make it clear what the output is. Some 
function definitions are shown at the end.

# Comparisons

```{r def_roll_funcs, echo = FALSE, message=FALSE}
#####
# Pure R version of package function
roll_regress_R <- function(X, y, width){
  n <- nrow(X)
  p <- ncol(X)
  out <- matrix(NA_real_, p, n)
  
  dummy_1 <- matrix(0.)
  dummy_2 <- numeric(p)
  dummy_3 <- numeric(p)

  is_first <- TRUE
  for(i in width:n){
    if(is_first){
      is_first <- FALSE
      qr. <- qr(X[1:width, ])
      R <- qr.R(qr.)

      # Use X^T
      X <- t(X)

      XtY <- drop(tcrossprod(y[1:width], X[, 1:width]))
    } else {
      x_new <- X[, i]
      x_old <- X[, i - width]

      # update R 
      rollRegres:::dchud_wrap(
        R, p, p, x_new, dummy_1, 0L, 0L, 0., 0., dummy_2, dummy_3)

      # downdate R
      rollRegres:::dchdd_wrap(
        R, p, p, x_old, dummy_1, 0L, 0L, 0., 0., dummy_2, dummy_3, 
        integer(1))

      # update XtY
      XtY <- XtY + y[i] * x_new - y[i - width] * x_old
    }

    coef. <- .Internal(backsolve(R, XtY, p, TRUE, TRUE))
    coef. <- .Internal(backsolve(R, coef., p, TRUE, FALSE))

    out[, i] <- coef.
  }

  t(out)
}

#####
# simple R version
roll_regress_R_for_loop <- function(X, y, width){
  n <- nrow(X)
  p <- ncol(X)
  out <- matrix(NA_real_, n, p)

  for(i in width:n){
    idx <- (i - width + 1L):i
    out[i, ] <- lm.fit(X[idx, , drop = FALSE], y[idx])$coefficients
  }

  out
}

#####
# zoo version
.zoo_inner <- function(Z) {
  fit <- lm.fit(Z[, -1, drop = FALSE], Z[, 1])
  fit$coefficients
}
library(zoo)
roll_regress_zoo <- function(x, y, width){
  Z <- cbind(y, X)
  rollapply(Z, width, FUN = .zoo_inner,  
            by.column = FALSE,  align = "right", fill = NA_real_)
}

#####
# roll_lm
library(roll)
roll_lm_func <- function(x, y ,width)
  roll_lm(X, matrix(y, ncol = 1), wdth, intercept = FALSE)$coefficients[, -1L]

# use one thread as other methods are easily to compute in parallel too
RcppParallel::setThreadOptions(numThreads = 1)

# compile functions
library(compiler)
roll_regress_R          <- cmpfun(roll_regress_R)
roll_regress_R_for_loop <- cmpfun(roll_regress_R_for_loop)
.zoo_inner              <- cmpfun(.zoo_inner)
roll_regress_zoo        <- cmpfun(roll_regress_zoo)
roll_lm_func            <- cmpfun(roll_lm_func)
```

We start by simulating the data

```{r sim_data}
set.seed(32981054)
n <- 10000
p <- 6
wdth = 50
X <- matrix(rnorm(p * n), n, p)
y <- drop(X %*% runif(p)) + rnorm(n)
df <- data.frame(y, X)
frm <- eval(parse(text = paste0(
  "formula(y ~ -1 + ", paste0("X", 1:p, collapse = " + "), ")")))
```

Then we check that the functions give the same (the function definitions are at
the end of this document)

```{r roll_gives_same}
library(rollRegres)
all.equal(roll_regress_R(X, y, wdth), roll_regres.fit(X, y, wdth)$coefs, 
          check.attributes = FALSE)
all.equal(roll_regress_R(X, y, wdth), roll_regres(frm, df, wdth)$coefs, 
          check.attributes = FALSE)
all.equal(roll_regress_R(X, y, wdth), roll_regress_zoo(X, y, wdth), 
          check.attributes = FALSE)
all.equal(roll_regress_R(X, y, wdth), roll_regress_R_for_loop(X, y, wdth), 
          check.attributes = FALSE)
all.equal(roll_regress_R(X, y, wdth), roll_lm_func(X, y, wdth), 
          check.attributes = FALSE)
```

and here then we compare the computation time

```{r benchmark_roll}
microbenchmark::microbenchmark(
  roll_regress            = roll_regres.fit(X, y, wdth),
  # this will be slower due to call to `model.matrix` and `model.frame`
  roll_regress_df         = roll_regres(frm, df, wdth),
  roll_regress_R          = roll_regress_R(X, y, wdth),
  roll_regress_zoo        = roll_regress_zoo(X, y, wdth),
  roll_regress_R_for_loop = roll_regress_R_for_loop(X, y, wdth),
  roll_lm = roll_lm_func(X, y, wdth),
  times = 5)

# here is the formula used above
frm
```

# Additional features

This section will cover some additional features.

## Expanding window

Here are expanding window regressions with additional output

```{r expanding_window}
#####
# simulate data
set.seed(65731482)
n <- 100
p <- 2
X <- matrix(rnorm(p * n), n, p)
y <- drop(X %*% runif(p)) + rnorm(n)

#####
# use package function
pck_out <- roll_regres.fit(
  X, y, width = 50L, do_downdates = FALSE, 
  do_compute = c("sigmas", "r.squareds", "1_step_forecasts"))

#####
# assign R-version
R_func <- function(X, y, width){
  n <- nrow(X)
  p <- ncol(X)
  out <- matrix(NA_real_, n, p)
  sigmas             <- rep(NA_real_, n)
  r.squared          <- rep(NA_real_, n)
  one_step_forecasts <- rep(NA_real_, n)

  for(i in width:n){
    idx <- 1:i
    fit <- lm(y[idx] ~ -1 + X[idx, , drop = FALSE])
    out[i, ] <- fit$coefficients

    su <- summary(fit)
    sigmas[i] <- su$sigma

    ss1 <- sum((y[idx] - mean(y[idx]))^2)
    ss2 <- sum(fit$residuals^2)
    r.squared[i] <- 1 - ss2 / ss1

    if(i < n){
      next_i <- i + 1L
      one_step_forecasts[next_i] <- fit$coefficients %*% X[next_i, ]
    }
  }

  list(coef = out, sigmas = sigmas, r.squared = r.squared,
       one_step_forecasts = one_step_forecasts)
}

R_out <- R_func(X, y, 50L)

#####
# gives the same
stopifnot(isTRUE(all.equal(R_out$coef              , pck_out$coefs)))
stopifnot(isTRUE(all.equal(R_out$sigmas            , pck_out$sigmas)))
stopifnot(isTRUE(all.equal(R_out$r.squared         , pck_out$r.squared)))
stopifnot(isTRUE(all.equal(R_out$one_step_forecasts, pck_out$one_step_forecasts)))
```

## Update in blocks
You can use the `grp` argument to make updates in blocks. E.g., here is an 
example with weekly data

```{r block_example}
#####
# simulate data
set.seed(68799379)
week <- as.integer(gl(25, 7))
head(week[1:10])
n <- length(week)
p <- 2
X <- matrix(rnorm(p * n), n, p)
y <- drop(X %*% runif(p)) + rnorm(n)

#####
# use package function
pck_out <- roll_regres.fit(
  X, y, width = 10L, grp = week, 
  do_compute = c("sigmas", "r.squareds", "1_step_forecasts"))

#####
# assign R-version
R_func <- function(X, y, width, grp){
  grp <- grp + 1L - min(grp)
  u_grp = unique(grp)
  n <- nrow(X)
  p <- ncol(X)
  out <- matrix(NA_real_, n, p)
  sigmas             <- rep(NA_real_, n)
  r.squared          <- rep(NA_real_, n)
  one_step_forecasts <- rep(NA_real_, n)

  start_val <- min(which(u_grp >= width))
  for(g in u_grp[start_val:length(u_grp)]){
    idx <- which(grp %in% (g - width + 1L):g)
    i <- which(grp == g)
    fit <- lm(y[idx] ~ -1 + X[idx, , drop = FALSE])
    out[i, ] <- sapply(fit$coefficients, rep, times = length(i))

    su <- summary(fit)
    sigmas[i] <- su$sigma

    ss1 <- sum((y[idx] - mean(y[idx]))^2)
    ss2 <- sum(fit$residuals^2)
    r.squared[i] <- 1 - ss2 / ss1

    if(g != max(grp)){
      next_g <- grp[min(which(grp > g))]
      next_g <- which(grp == next_g)
      one_step_forecasts[next_g] <- X[next_g, ] %*% fit$coefficients
    }
  }

  list(coef = out, sigmas = sigmas, r.squared = r.squared,
       one_step_forecasts = one_step_forecasts)
}

R_out <- R_func(X, y, 10L, grp = week)

#####
# gives the same
stopifnot(isTRUE(all.equal(R_out$coef              , pck_out$coefs)))
stopifnot(isTRUE(all.equal(R_out$sigmas            , pck_out$sigmas)))
stopifnot(isTRUE(all.equal(R_out$r.squared         , pck_out$r.squared)))
stopifnot(isTRUE(all.equal(R_out$one_step_forecasts, pck_out$one_step_forecasts)))
```

## Block with minimum required number of observation
Suppose that we are performing rolling regression on stock data over a yearly 
window. Further, suppose that there are som gaps in the data where we do not 
have data and we require at least 6 months of data. This can be done as follows

```{r min_obs}
#####
# simulate data
set.seed(96235555)
n   <- 10L * 12L * 21L # x years w/ 12 months of 21 trading days
month <- (seq_len(n) - 1L) %/% 21L + 1L # group by months
set.seed(29478439)
X <- matrix(rnorm(n * 4L), ncol = 4L)
y <- rnorm(n)

# randomly drop rows
keep <- seq_along(y) %in% sample.int(nrow(X), as.integer(n * .5))
X     <- X    [keep, ]
y     <- y    [keep]
month <- month[keep]

#####
# use package function
pck_out <- roll_regres.fit(
  X, y, width = 12L, grp = month, min_obs = 21L * 6L, do_downdates = TRUE, 
  do_compute = c("sigmas", "r.squareds"))

#####
# assign R-version
R_func <- function(X, y, width, grp, downdate, min_obs){
  grp <- grp + 1L - min(grp)
  u_grp = unique(grp)
  n <- nrow(X)
  p <- ncol(X)
  out <- matrix(NA_real_, n, p)
  sigmas    <- rep(NA_real_, n)
  r.squared <- rep(NA_real_, n)

  start_val <- min(which(u_grp >= width))
  for(g in u_grp[start_val:length(u_grp)]){
    idx <-
      if(downdate)
        which(grp %in% (g - width + 1L):g) else
          which(grp %in% 1:g)
    i <- which(grp == g)
    if(length(idx) < min_obs)
      next
    fit <- lm(y[idx] ~ -1 + X[idx, , drop = FALSE])
    out[i, ] <- sapply(fit$coefficients, rep, times = length(i))

    su <- summary(fit)
    sigmas[i] <- su$sigma

    ss1 <- sum((y[idx] - mean(y[idx]))^2)
    ss2 <- sum(fit$residuals^2)
    r.squared[i] <- 1 - ss2 / ss1
  }

  list(coef = out, sigmas = sigmas, r.squared = r.squared)
}

R_out <- R_func(X, y, width = 12L, downdate = TRUE, grp = month, 
                min_obs = 21L * 6L)

#####
# gives the same
stopifnot(isTRUE(all.equal(R_out$coef     , pck_out$coefs)))
stopifnot(isTRUE(all.equal(R_out$sigmas   , pck_out$sigmas)))
stopifnot(isTRUE(all.equal(R_out$r.squared, pck_out$r.squared)))
```

# Session info

```{r ses_info}
sessionInfo()
```

# Function definitions

Here are the function definitions

```{r def_roll_funcs, eval = FALSE}
```
